### 2011-03-14 trunk/lids/twittersearchwrap/paper/paper.tex

\documentclass{llncs}
%\documentclass[a4paper]{article}

%\usepackage{fullpage}
\usepackage{setspace}
\usepackage{url}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{latexsym}

\usepackage{graphicx}
\usepackage{subfig}

%\onehalfspacing

%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
%\newtheorem{example}{Example}
\newtheorem{query}{Query}

\newcommand{\nop}[1]{}

\setlength{\tabcolsep}{2pt}

\newcommand{\myremark}[2]{{\textbf{\marginpar{$\parallel$}(* \textit{#1:} #2 *)}}}
%\newcommand*\justify{%
%  \fontdimen2\font=0.4em% interword space
%  \fontdimen3\font=0.2em% interword stretch
%  \fontdimen4\font=0.1em% interword shrink
%  \fontdimen7\font=0.1em% extra space
%}

\title{Mapping Microblog Posts to Encyclopedia Articles}
\author{Uta L\"{o}sch \and David M\"{u}ller \and Andreas Harth}
\institute{
	Karlsruhe Institute of Technology (KIT), D-76131 Karlsruhe, Germany\\ 
	\email{uta.loesch@kit.edu},\\
	\email{david.mueller@student.kit.edu},\\
	\email{harth@kit.edu}
}
\begin{document}

\maketitle

\begin{abstract}
To facilitate understanding the meaning of a hash tag in the context of Twitter, we have developed a system which annotates hash tags based on Twitter search results with encyclopedia entities that best describe the hash tag, thus offering means of quickly grasping the meaning of the hash tag and at the same time providing starting points for further exploration of the hash tags' context. In an evaluation we show that the encyclopedia entities used for annotation represent the content in a suitable manner and that the annotations remain stable over time, i.e. when executing the same search at different times, the same encyclopedia entities are returned.
\end{abstract}

Keywords: Twitter, Wikipedia, DBpedia, RDF

\nop{
http://tagdef.com/ has a list of popular tags
}

\section{Introduction}

Microblogging has become more and more popular over the last years, especially with the most prominent and most widely used microblogging platform Twitter\footnote{\url{http://twitter.com}} which has become very influential over the last years. The idea of micro-blogs is the same as that of blogs, except that the message length is restricted, in the case of Twitter the restriction is 140 characters. Twitter has a total of about 190 million of users who produce 65 million Twitter messages a day. Thus, Twitter represents a huge data source on the web.

On microblogging platforms it is possible to search for messages containing a specific hash tag. This search returns a fixed number of most recent messages containing the tag. However, it is hard to grasp the context of the results and to get further information on the topic that was searched for. As each single microblog post is very short and contains little information, it is necessary to parse the whole set in order to get an overview of the context(s) in which the search term is used.

To facilitate this putting into context of the search results, we propose to annotate search results with a set of entities which reflect the content of the result feeds. These entities will not only help to understand the search terms, but also serve as a starting point for refining the search or searching for further information related to the search result.

Furthermore, this annotation will help to understand hash tags. Hash tags are frequently used to associate messages with a specific topic, place, person or event on microblogging platforms. For example, messages talking about what is going on at the European Semantic Web Conference 2011 will frequently be tagged with $\#eswc2011$. If a user encounters this hash tag and does not understand it, he can search for it on the microblogging platform. However, as discussed before, to understand a search result is difficult and time-consuming. Using our approach the search will not only return messages using the tag but also a set of entities which have been detected in the search result. These entities will help the user to gain an understanding in which context the hash tag is used and what it may mean. If a hash tag can be associated with an entity with high confidence, the entity can be seen as descriptive for the hash tag. A system may be envisioned which automatically finds entitities which are equivalent to a hash tag. Thus, these hash tag-entity pairs may be used similarly to systems in which hash tags can be defined manually, such as is possible on \#tagdef\footnote{\url{http://www.tagdef.com}}.

Finding entities matching a query, especially a query for a given hash tag is difficult due to synonyms and homonyms in the set of hash tags that are frequently used. For example, both the hash tags $\#ka$ and $\#karlsruhe$ are frequently used when talking about topics which are related to the city Karlsruhe. The tag $\#ka$ is however also used in the context of the car type Ford Ka. Thus, when searching for $\#ka$ on a microblogging platform, messages relating to Karlsruhe will be found as well as messages related to Ford Ka. An approach which aims at finding relevant entities for a given search will have to be able to deal with these problems.

In this paper, we present a system which automatically annotates a microblog search result with encyclopedia entities. While it would be more interesting to annotate each single microblog message with relevant entities, microblog messages are too short to contain much relevant information which could be used as input for the annotation tool. Thus, we chose to generate annotations for the whole search result.

The implementation we provide is based on Twitter, as it is the most widely used and biggest microblogging platform, and on Wikipedia. The motivation for choosing Wikipedia as entity source was that it offers a wide coverage of topics. Furthermore, automatic tools for annotating text with Wikipedia entities are readily available (see \cite{key:wikifier}). Note however that a direct mapping of hash tags to entities is not possible in many cases: e.g. there is no entity in dbpedia which describes the European Semantic Web Conference 2011, thus the tag $\#eswc2011$ can not mapped directly.

Annotations should be stable over time, unless the topics discussed in the microblog messages containing the search term change. In an evaluation of our approach we have tried to analyze the stability of the result.

Our contributions thus are:
\begin{itemize}
	\item an approach for finding entities which are related to a search on microblogging platforms
	\item a formatting of microblogging messages as RDF
	\item an implementation of our approach
	\item an evaluation of the stability and relevance of the entities found with our approach.
\end{itemize}

The remainder of this paper is organized as follows: in Section~\ref{sect:relWork} we discuss related work, in Section~\ref{sect:method} we present our approach, the results of our evaluation are presented in Section~\ref{sect:eval}, before we conclude in Section~\ref{sect:conclusion}.

\section{Related Work}
\label{sect:relWork}

Bringing Semantic Web technologies and Semantic Web technologies together, has been proposed several times.

The SemanticTweet\footnote{\url{http://semantictweet.com}} service is a tool which allows for the generation of a FOAF file describing one's network of followers and friends on Twitter. Thus, an automatic way of mapping the Twitter social network to semantic data is made available.

Passant et al. \cite{key:smob} have proposed a data model for making Twitter data available on the Semantic Web. They propose a data model which allows for the association of URIs with users, microblogs and microposts. To this end, SIOC and FOAF vocabularies are used and extended. Specifically, the new concepts \emph{Microblog} and \emph{MicroBlogPost} are introduced. Additionally, they propose the use of so-called semantic hash tags. The idea is to use URIs as hash tags (e.g. \emph{\#geo:Paris\_France}). It becomes thus possible to link microposts to entities in the Linked Open Data cloud via these semantic hash tags. 
In contrast to this approach we do not require users to change their behavior by requiring them to use another type of hash tags. Instead, our method finds Linked Open Data entities which are related to the query automatically.

Softic et al \cite{key:softic} have developed a framework and system for mining data from social networks. They have instantiated their system to analyze data from Twitter. In their system, they collect data from one or several users. The system by Passant et al. \cite{key:smob} is used for transforming twitter data to semantic data. The triplified data can then be analysed using SPARQL. They propose that the result data could be enriched using entities from other Linked Open Data vocabularies like dbpedia\footnote{\url{http://dbpedia.org}} or GeoNames\footnote{\url{http://www.geonames.org}}, but they do not provide a method for achieving this.

Relating semantic concepts with twitter messages has been attempted by various authors:
Stankovic et al. \cite{key:stanko} propose to use the Zemanta\footnote{\url{http://developer.zemanta.com/}} keyword extraction API to detect topics of single Twitter messages. 
% from stakovic et al.
%In order to make tweets more searchable and make them useful for the described
%scenario one would need to have represent them in a structured format and associate
%those tweets with the topics that refer to. As a basic approach we process the text of
%the tweets using Zemanta16 keyword extraction API. This approach gives DBPedia
%concepts related to a tweet, but is useful only on a limited number of tweets directly
%mentioning a particular topic. In the example data set we produced from ESWC2010,
%tweet topics are represented using dc:subject property.
Instead of mining single messages which provide little signal to the extraction algorithm we use an aggregation of messages (as proposed by Wagner \cite{key:clauwa}) as input to the extraction component.

Wagner \cite{key:clauwa} proposes to use aggregations of social awareness streams for knowledge extraction. As such, our approach is closely related.

Kinsella et al. \cite{key:kinsella} use hyperlinks in posts to derive a richer representation of those posts. We too use external data to augment the set of Twitter messages supplied to the Wikifier system.

The challenge task in the WePS-3 Online Reputation Management Task\cite{key:clef} requires systems to detect mentions of company names in Twitter messages.
The corpus contains evaluation results solicited via Mechanical Turk. In contrast to the WePS-3 task, we associate generic hash tags to Wikipedia/DBpedia concepts.

\section{Method Overview}
\label{sect:method}

The goal of our method is to, given a search query $q$ as input, provide the user with an RDF\footnote{\url{http://www.w3.org/RDF/}} document describing the search result $R$, i.e. the most recent messages that match the query, the authors of these messages and the most relevant entities $E$ for this result.

The architecture of our proposed system is shown in Figure~\ref{fig:arch}. Upon
receipt of a query $q$ four steps are performed. In the following we will illustrate
these steps using the example of a search for the hash tag $\#Libya$, a hashtag
which is frequently used to describe posts related to the country of Libya.

\begin{figure}[htb]
  \centering
  \includegraphics[width=.7\linewidth]{architecture}
  \caption{System Architecture}
  \label{fig:arch}
\end{figure}

\begin{enumerate}
	\item In a first step, the most recent search results $R$ for the query are
	obtained.\newline In our example, calling the Wrapper with query $\#Libya$ will
	return a document containing all most recent Twitter posts whose content
	matched the search string $\#Libya$ and their meta information. 
	\newline
	An example\footnote{\url{http://twitter.com/SenJohnMcCain/statuses/46619460060196865}}
for a single retrieved post is the following:
\begin{verbatim}
"Arab League calls for no-fly zone in \#Libya 
http://t.co/ZsLNIWa"}}
\end{verbatim}
,posted by Twitter user \texttt{SenJohnMcCain}; geo coordinates\footnote{The original Twitter message posted by SenJohnMcCain was not geotagged} and the date the message has been posted are furthermore known.

  \item In a second (optional) step hyperlinks posted in messages of the fetched
  feed are followed, content of the referenced website is fetched and processed by the Wrapper. The idea is to acquire additional content which is used as additional signal when searching for entities. We have thought of other methods for enriching the result, like searching for other messages posted by users which have posted messages in the result set. However, these alternative methods are more likely to introduce noise in the signal which is used for obtaining relevant entities.

In our example, the message contains a link to
\url{http://t.co/ZsLNIWa}, which redirects to an article on the situation in
Libya published on \url{http://reuters.com}. Dereferencing this URL and grabbing
the text found at the link should return the following text:
\begin{verbatim}
"Arab League calls for Libya no-fly zone-state TV - CAIRO, 
March 12 (Reuters) - The Arab League on Saturday called on the
U.N. Security Council to impose a no-fly zone on Libya, 
Egyptian state television reported, a decision that would give
a regional seal of approval that NATO has said is needed for 
any military action."
\end{verbatim}
	\item In a third step entities matching the search result are retrieved. The
	content of the Twitter messages that matched the search query and, if available, the content of referenced websites is merged to a single input string and used as input for the annotation. The result of this step is a list of matching articles of the English Wikipedia.

In our example, the contents of all posts related to the search term '\#Libya'
are merged with the retrieved external website content. Thus, the post obtained from
step 1 would be matched with the website content obtained in step 2 into a
single input stream:
\begin{verbatim}"Arab League calls for no-fly zone in #Libya
Arab League calls for Libya no-fly zone-state TV - CAIRO, 
March 12 (Reuters) - The Arab League on Saturday called on the 
U.N. Security Council to impose a no-fly zone on Libya, 
Egyptian state television reported, a decision that would give
a regional seal of approval that NATO has said is needed for 
any military action."}
\end{verbatim}
This string is appended to the Twitter content and the external content of all other posts in the result.

In the example, we would expect to find entities which are related to Libya,
like Libya, Gaddafi or the Arab\_League.

	\item	Finally, an RDF document is generated representing the search results and
	the annotations. Continuing our example, the output RDF document should contain the following
	triples to reference the DBPedia mappings:
	\begin{verbatim}
	<http://km.aifb.kit.edu/services/twittersearchwrap/searchwrap
	?q=\%23lybia>
     rdfs:seeAlso <http://dbpedia.org/resource/Lybia> ,
        <http://dbpedia.org/resource/Gaddafi> ,
        <http://dbpedia.org/resource/Arab\_League> .
  \end{verbatim}
	The Twitter posts matching the search term \#Libya are likewise all described
	in the output document. The description for the post from our example 
	%abbreviated \texttt{m} 
	results in the following triples:
	\begin{verbatim}
	<http://km.aifb.kit.edu/services/twitterwrap/statuses/
	show/46619460060196865\#id> 
		foaf:page <http://twitter.com/SenJohnMcCain/statuses/
			46619460060196865>;
		dc:date "2011-03-12T17:12:00Z" ;\newline geo:lat "49.010239" ;
		geo:long "8.411879" ;
		foaf:maker <http://km.aifb.kit.edu/services/twitterwrap/users/
			show?screen\_name=SenJohnMcCain\#id> ;
		dc:description "Arab League calls for no-fly zone in \#Libya 
			http://\newline t.co/ZsLNIWa" . 
	\end{verbatim}

\end{enumerate}

\section{Implementation}

In the Twitter Search Wrapper\footnote{Online at \url{http://km.aifb.kit.edu/services/twittersearchwrap/}} we provide an implementation of our approach.

The implementation is based on a set of existing components: 
\begin{itemize}
\item Twitter search results are obtained by using the Twitter search API\footnote{\url{http://dev.twitter.com/doc/get/search}}, a RESTful API which delivers search results in JSON. The result is a feed containing the 100 most recently published Twitter messages, which are publicly visible and match the query, and their authors. The messages themselves are described by their content, publishing date, authors
and optionally a geographic location.

\item The optional step of dereferencing URLs which are posted in messages contained
in the result set can be triggered using the option \texttt{extern=true} in our
system. To keep the input size for the next steps manageable, we do not use the
whole content of the referenced sites but only the first 50 words. The content
retrieved from external websites is beautified for our purpose removing
HTML-tags and scripting elements using the CyberNeko Java HTML Parser
Library\footnote{\url{http://sourceforge.net/projects/nekohtml/}}.

\item The Wikifier \cite{key:wikifier} is used for obtaining content annotations. This tool takes a text as input and returns a set of Wikipedia entities which are relevant for the analyzed text. It represents a state-of-the-art annotation tool. It was chosen as it allows for finding entities not from some tool-specific knowledge base but from dbpedia, which is one of the most widely used data sources within the Linked Open Data cloud.

\item The system generates and outputs an RDF document in RDF/XML Syntax containing a
description of the retrieved messages matching the search query for the given query and references to the DBpedia articles that have been returned by the Wikifier. It describes
the result by means of the popular vocabularies FOAF \cite{key:foaf}, Dublin
Core\footnote{\url{http://dublincore.org/documents/dcmi-terms/}}, the Basic Geo
Vocabulary \cite{key:geo}. A single twitter message is described by its
individual page using \texttt{foaf:page} predicate. The date and time the
message has been posted is specified using \texttt{dc:date} predicate. The messages content
is annotated using \texttt{dc:description}, while the geographic location of the
message is indicated by \texttt{geo:lat} and \texttt{geo:long} predicates. The
creator of the Twitter message is referenced by \texttt{foaf:maker} predicate which points to a service that
uses RDF to describe information related to a Twitter user which has previously
been implemented here at the institute as the message creators URI.

\end{itemize}

The annotations are added to the search result using \texttt{rdf:seeAlso} links to DBpedia entities \cite{key:dbpedia} which are obtained by a direct transformation of the Wikipedia entities' URLs. DBpedia has been chosen as these entities allow for a more coherent presentation of the content in the context of RDF.

%The example twitter message given in the previous section is represented in
%the output of the Twitter Search Wrapper as follows:\newline
%\linebreak
%\texttt{<rdf:Description
%rdf:about="http://km.aifb.kit.edu/services/\newline
%twitterwrap/statuses/show/1234567890\#id">\newline
%	<foaf:page
%	rdf:resource="http://twitter.com/sample\_user\_12345/\newline
%	statuses/1234567890"/>\newline
%	<dc:date>2011-02-23T12:00:00Z</dc:date>\newline
%	<geo:lat>49.010239</geo:lat>\newline <geo:long>8.411879</geo:long>\newline
%	<foaf:maker
%	rdf:resource="http://km.aifb.kit.edu/services/\newline
%	twitterwrap/users/show?screen\_name=sample\_user\_12345\#id"/>\newline
%	<dc:description>\newline Just returned from a wonderful holiday in \#ka.
%	\url{http://www.karlsruhe.de/stadt/tourismus.en} @KarlsruheTweets\newline
%	</dc:description>\newline
%	</rdf:Description>\newline}
%	\linebreak
%Furthermore the mappings to DBpedia articles returned by the Wikifier
%are represented in the output document. In the Twitter Search Wrapper the
%representation is done the following way, again using the introductory
%example from the previous section:\newline
%\linebreak
%	\texttt{<rdf:Description rdf:about=""/>}\newline
%	\texttt{<rdfs:seeAlso
%	rdf:resource="http://dbpedia.org/resource/Karlsruhe"/>}\newline
%	\texttt{<rdfs:seeAlso
%	rdf:resource="http://dbpedia.org/resource/Germany"/>}\newline
%	\texttt{<rdfs:seeAlso rdf:resource="http://dbpedia.org/resource/\newline
%	Karlsruhe\_Institute\_of\_Technology"/>}\newline
%	\texttt{</rdf:Description>}\newline

\section{Experiments and Evaluation}
\label{sect:eval}

We have evaluated our approach to assess whether the entities which are found as related to a specific hash tag are useful and stable over time.

\subsection{Setup}

We have conducted experiments on a set of queries which were identified to be hash tags which are frequently used on Twitter.
To obtain these, we have sampled the Twitter message feed using Twitter's
Streaming API for one minute twice an hour from 2011-03-03 to 2011-03-06. The
top-50 hashtags dervied from the aggregate sample are listed in Table
\ref{tbl:hashtags}, count is the number of times we encountered them during the
collection period.

\begin{table}[ht*]
\centering
\begin{tabular}{ l|l|l|l|l|l|l|l|l }
Hashtag & Count & $N_{ext}$ & $E(aut)$ & $Var(aut)$ & $E(ts)$ & $Var(ts)$ & $E(ts_{ext})$ & $Var(ts_{ext})$\\
\hline
ff &     460 & 48 & 2.2936 & 0.7014 & 0.1268 & 0.0681 & 0.2037 & 0.0281 \\
np &     172 & 48 & 0.4737 & 0.0287 & 0.0655 & 0.0196 & 0.0717 & 0.0195\\
nowplaying & 144 & 17 & \textbf{0.4514} & \textbf{0.0251} & 0.0565 & 0.0198 & 0.0458 & 0.0212 \\
winning &     134 & 48 & 0.8236 & 0.2365 & 0.1101 & 0.0666 & 0.1816 & 0.0272 \\
fb & 105 & 1 & 1.1705 & 1.1826 & 0.1083 & 0.0933 & -- & -- \\
rt & 83 & 48 & 1.7980 & 3.7309 & 0.0979 & 0.0763 & 0.1487 & 0.0267 \\
mentionke & 61 & 48 & 7.5459 & 195.6829 & 0.1406 & 0.0921 & 0.2215 & 0.0408 \\
blackpeoplemovies & 57 & 3 & 90.6195 & 2007.5551 & 0.3327 & 0.0606 & 0.0349 & 0.0244\\
tigerblood & 56 & 48 & 3.4029 & 4.8740 & 0.1497 & 0.0692 & 0.2330 & 0.0278 \\
thataintwinning & 53 & 48 & 16.8075 & 236.8694 & 0.0768 & 0.0435 & 0.1668 & 0.0387 \\
fail & 53 & 48 & 2.3356 & 2.0669 & 0.1927 & 0.1517 & 0.1073 & 0.0460 \\
teamfollowback & 52 & 48 & 0.9269 & 0.4871 & 0.1912 & 0.0932 & 0.1813 & 0.0245 \\
iftwitterwashighschool & 51 & 48 & 114.8342 & 1771.4564 & 0.5113 & 0.0405 & 0.5144 & 0.0352 \\
bieberfact & 51 & 48 & 11.0126 & 63.0455 & 0.0587 & 0.0290 & 0.0765 & 0.0270 \\
news & 46 & 0 & 1.8696 & 0.1943 & 0.1024 & 0.0210 & -- & --  \\
jobs & 45 & 0 & 1.3866 & 0.4198 & 0.1006 & 0.0192 & -- & -- \\
libya & 44 & 1 & 16.0309 & 117.2024 & 0.2701 & 0.0427 & -- & -- \\
followmejp & 44 & 0	& 1.5398 & 0.1255 & 0.0688 & 0.0410 & -- & --\\
partiu & 41 & 48 & 7.5809 & 84.0531 & 0.6042 & 0.2391 & 0.5313 & 0.2438 \\
sougofollow & 39 & 0 & 2.1010 & 0.1820 & 0.0720 & 0.0427 & -- & -- \\
damnitstrue & 39 & 48 & 5.7507 & 16.5725 & 0.1805 & 0.1387 & 0.2177 & 0.1579 \\
tilltheworldends & 37 & 48 & 85.4526 & 1366.8588 & 0.5682 & 0.0709 & 0.5116 & 0.0324 \\
soalcinta & 35 & 48 & 8.4143 & 103.7817 & 0.3451 & 0.1987 & 0.3120 & 0.2014 \\
nicovideo & 35 & 0 & 51.0720 & 5521.3529 & 0.3604 & 0.1893 & -- & -- \\
quran & 33 & 48 & 45.8290 & 90.4157 & 0.4271 & 0.0501 & 0.4188 & 0.0387 \\
%\end{tabular}
%\begin{tabular}{ l|l|l|l|l|l|l|l|l }
%Hashtag & Count  \\
%\hline
nw & 32 & 48 & 2.6054 & 2.1094 & 0.0672 & 0.0231 & 0.0901 & 0.0418 \\
quote & 31 & 48 & 4.3007 & 0.5706 & 0.1074 & 0.0197 & 0.1012 & 0.0212 \\
shoutout & 30 & 48 & 2.6030 & 3.3256 & 0.2347 & 0.0987 & 0.2526 & 0.0488 \\
jfb & 30 & 48 & 23.9911 & 1453.3113 & 0.2483 & 0.1654 & 0.2158 & 0.1379 \\
carnaval & 27 & 48 & 34.2574 & 1100.9818 & 0.3229 & 0.2134 & 0.2578 & 0.1646 \\
tfb & 26 & 48 & 2.5994 & 0.8777 & 0.2292 & 0.1766 & 0.1760 & 0.0319 \\
followme & 26 & 0 & 3.2194 & 0.4650 & 0.1209 & 0.0952 &	-- &	-- \\
followfriday & 25 & 48 & 13.6749 & 20.0701 & 0.1910 & 0.1152 & 0.2060 & 0.0676 \\
follow & 24 & 48 & 3.0945 & 1.9622 & 0.2916 & 0.1927 & 0.2046 & 0.0305 \\
egypt & 23 & 16 & 3.1850 & 3.2745 & 0.1357 & 0.0239 & 0.0704 & 0.0263 \\
retweet & 22 & 48 & 4.6991 & 13.4434 & 0.1570 & 0.1232 & 0.1353 & 0.0306 \\
fact & 22 & 48 & 5.5613 & 12.0766 & 0.0915 & 0.0762 & 0.0783 & 0.0612 \\
win & 20 & 0 & 4.2949 & 5.1856 & \textbf{0.0347} & 0.0221 & -- & -- \\
tweetmyjobs & 20 & 0 & 35.5697 & 1590.6997 & 0.3244 & 0.0621 & -- & -- \\
music & 20 & 1 & 4.4665 & 4.6392 & 0.0833 & 0.0196 & -- & -- \\
in & 20 & 3	& 8.2367 & 6.8944 & 0.0649 & 0.0236 & \textbf{0.0251} & 0.0206 \\
imagine & 20 & 48	& 8.3378 & 30.4317 & 0.2465 & 0.1759 & 0.1813 & 0.1375 \\
nowwatching & 19 & 48 & 7.9110 & 8.6833 & 0.0672 & 0.0214 & 0.1293 & 0.0211 \\
cumannanyaindonesia & 19 & 48 & 77.9680 & 2739.7782 & 0.3901 & 0.1885 & 0.4323 & 0.2033 \\
twexit & 18 & 48 & 12.5672 & 255.8178 & 0.3125 & 0.2148 & 0.2417 & 0.1783 \\
lcv & 18 & 48 & 28.6309 & 477.9193 & 0.4236 & 0.2395 & 0.2813 & 0.1969 \\
fui & 18 & 48 & 11.5515 & 165.3850 & 0.4896 & \textbf{0.2447} & 0.4635 & \textbf{0.2448} \\
bot &      18 & 1 & 13.3533 &	8.2469 & 0.2708 & 0.1975 & -- & --  \\
nemuritsuzuketeshinu & 17 & 48 & \textbf{314.1120} & \textbf{25053.5657} & \textbf{1.0000} & \textbf{0.0000} & \textbf{1.0000} & \textbf{0.0000} \\
fato & 17 & 48 & 17.6223 & 569.5627 & 0.1875 & 0.1523 & 0.2353 & 0.1304 \\
\end{tabular}
\caption{Popular hashtags sampled during a three-day period.}\label{tbl:hashtags}
\end{table}

The goal of our evaluation was to examine the stability of the entity annotations over time. To achieve this, we have repeatedly executed the same set of queries and analysed the returned data.

In principle, two methods for data acquisition were identified: \begin{definition}[Data Acquisition Methods]
\begin{itemize}
\item {\bf Temporal Sliding Window}\newline The query is executed and the n
most recent results are retrieved every t minutes. The Temporal Sliding Window
method may lead to duplicate messages in the retrieved documents, but needs to
handle much less data and allows for an analysis over larger time frames.
\item {\bf $k$-item Sliding Window}\newline  The k-item Sliding Window methods
retrieves all messages that match the query exactly one time. Each time k new messages are
available matching the given query a document containing those k messages
is retrieved. This method can lead to high data volume and needs constant
monitoring of recently posted messages matching the search term.
\end{itemize}
\end{definition}

The disadvantage of the first method is that it is neither possible to guarantee that all messages matching the search are collected, nor is it possible to avoid overlaps between the search results. However, while the second method seems to overcome these problems, it does not: Using the $k$-item sliding window method, it is not clear when the next search result should be fetched: even when frequently querying for new messages, if updates are frequent it is likely to lack messages in the result set.

For both data set queries were executed twice per hour using the Temporal
Sliding Window method for data acquisition in a 24 hour time frame. Data for the
trending topics was collected on 2011-02-24, for the frequent hash tags data was
collected on 2011-03-07\footnote{Collected data is available under
\url{http://people.aifb.kit.edu/aha/2011/twittersearch/twittersearch-2011-03-10.zip}}.

\subsection{Measures}

For analysing the collected data, several measures were defined:

\begin{definition}[Evaluation Measures]
Given disjoint result sets $R_1,\ldots,R_n$ for a query $q$ obtained at times $t_1,\ldots,t_n$ and annotated with sets of entities $E_1,\ldots,E_n$, where each result set $R_i$ consists of twitter messages $r_i1,\ldots,r_iK$ with time stamps $t_i1,\ldots,t_iK$ we define:
\begin{itemize}
	\item The average update time within a result set $aut(R_i)$ for a query is the average time that lies between messages that match the query which produced the search result: $$aut(R_i)=\frac{\sum_{j=1}^{n}t_{i,j+1}-t_{i,j}}{K-1}$$
	A high $aut(R)$ means that new messages that match the query are not generated very often. If new messages matching the query are frequently published, then the $adt(R)$ is low.

	The mean $aut(q)$ and the variance $autv(q)$ of the $aut(R_i)$ can be used to describe update properties for the result sets that are generated for a given query. E.g. if for a given query $aut(q)$ is high and $autv(q)$ is low, then the rate at which messages are published is stable at a quite low level.
%	\item The global stability $gs$ of the results shows how stable the whole set of entity terms for a query is. It is defined as: $$gs(q)=\frac{|\bigcap_{i=1}^{n}E_i|}{|\bigcup_{i=1}^{n}E_i|}$$
%	A value of $gs(q)$ of $1$ means that the set of entities is the same at each point in time $t_i$. On the other hand, a value of $0$ means that at least two of the sets of entities, with which the results were annotated, are disjoint.
	\item The stability for a single term is defined as the fraction of result sets for a query $q$ in which it occurs: 
$$sts(q,t)=\frac{|\{i|t\in E_i\}|}{n}$$
	If the value of $sts$ is $1$ the term occurs in all of the result sets for the query, if it is close to $0$ it occurs in few result sets.
	\item The time stability $ts$ for a query $q$ at time $t_i$ is the fraction of
	entities which are unchanged between the result set at $t_{i-1}$ and $t_i$: 
	$$ts(q,t_i)=\begin{cases} 1 & i=1 \\
														1 & E_{i-1}\cup E_i = \emptyset\\														
														\frac{|E_{i-1}\cap E_i|}{|E_{i-1}\cup E_i|} & i>1\end{cases}$$
	A value of $1$ means that the two sets of entities are identical, a value of $0$ means they are disjoint.

	The variance of time stabilities $ts(q,t_i)$ over all $t_i$ helps to determine whether the rate at which the found entities change is rather stable or whether it is changing a lot over time.
\end{itemize}
\end{definition}

\subsection{Results}

An overview of the results we obtained for the frequently used hash tags is given in Table~\ref{tbl:hashtags}. The most frequent entities for some of the tags are listed in Table~\ref{tbl:entities}. In total, we have analysed 48 result sets for each of the query tags. However, in some cases, the analysis which includes the content of external web pages led to a timeout of the Wikifier. $N_{ext}$ gives the number of result sets which we could analyse when the external content is included. $E(aut)$ and $Var(aut)$ give the results for analysing the average update times of the result sets. A low $E(aut)$ means that updates are frequently posted, a high $E(aut)$ means that updates are rather infrequent. In the case of our analysis estimated update frequencies range from $0.45$ seconds to $314.11$ seconds, i.e. approximately $5$ minutes. In order to avoid any duplicates in the result sets, an average update time of less than $18$ seconds is needed. Note that update times below this value do not guarantee disjointness of the result sets, as the value is only an average, and high update frequencies in one result set may outweigh low update frequencies in other result sets. Thus, we have additionally analysed the variance of the $aut$s. A low variance means that the update intervals remain relatively stable over time, while a high variance indicates that messages using the hash tag were posted rather irregularly. In general, higher average update times are associated with a higher variance in the update times. In some cases however, the variance of the average update time is surprisingly low when the mean average update time is rather high: this is e.g. the case for the tags $\#bieberfact$, $\#quran$, $\#followfriday$. For these tags updates came at a relatively regular level.

\begin{table}[ht]
\centering
\begin{tabular}{l|l|l|l}
Tag & extern & $E$ & $sts$ \\%& $E_2$ & $sts(E_2)$ & $E_3$ & $sts(E_3)$ \\
\hline
$\#tilltheworldends$ & false & Britney Spears .9583 \\
& &  iTunes & .6875 \\
& &  itunes & .625 \\ 
$\#nowplaying$ & false & iTunes & .5208 \\
& & Michael Jackson & .333 \\
& & Lady Gaga & .333 \\
$\#nowplaying$ & true & YouTube & .8235 \\
& & Hip Hop Music & .6471 \\
& & The Notorious B.I.G. & .5883 \\
$\#np$ & false & Chris Brown\_(entertainer) & .5 \\
& & Usher & .3125 \\
& & Lil Wayne & .2917 \\
$\#np$ & true & Chris Brown\_(entertainer) & .5 \\
& & YouTube & .375 \\
& & Ne-Yo & .315 \\
$\#thataintwinning$ & false & Mozilla Firefox & .2291 \\
& & Twitter & .2291 \\
& & twitter & .1667 \\
$\#thataintwinning$ & true & twitter & .5208 \\
& & Twitter & .5 \\
& & Two and a half men & .125 \\
$\#blackpeoplemovies$ & false & Devil & .4791 \\
& & Tyler Perry & .2708 \\
& & Harry Potter & .2292 \\
$\#blackpeoplemovies$ & true & Amazon.com & 1 \\
& & YouTube & 1 \\
& & amazon.com & 1
\end{tabular}
\label{tbl:entities}
\caption{Most frequent entities and their single term stabilities}
\end{table}

We also report the average time stability over all time points as well as the variance for each hash tag. A high time stability indicates that more or less the same entities were used for annotation at all times. The highest stabilities were achieved for the tags $\#nemuritsuzuketeshinu$ and $\#partiu$, because no (resp. very few) entities could be associated with the results of this search. This can be explained as the tags are only used in Japanese resp. Spanish twitter messages. However, the Wikifier analyses the content of the results as being English text. A solution could be to search only for English messages, however, this is not possible as Twitter messages have no language tags and location attributes are no sufficient indicator for the language in which a user will write messages.

The third highest stability was achieved by the tag $\#tilltheworldends$ which refers to Britney Spears' newest single. The different URIs which are used for annotation are explained by the fact, that the Wikifier does not return normalized entities.

Very low time stability was obtained for the tag $\#nowplaying$ which is used to tell which song one is currently listening to (a similarly low stability is obtained for the tag $\#np$ which is often used synonymously with $\#nowplaying$). As these messages contain all kinds of artist names and songs names, it is thus expected that the entities which are found vary between the result sets.

The effect of considering external content in the annotation is beneficial in some cases, in others the effect was destabilizing. The highest decrease in stability was observed for the tag $\#thataintwinning$, the highest increase occured for the tag $\#blackpeoplemovies$. Our hypothesis is that destabilization occurs especially in cases where the posted links cover a variety of topics, while for more specific hash tags referenced web sites are more likely to cover a specific topics and will thus be stabilizing.

\section{Conclusion}
\label{sect:conclusion}

We have presented an approach for annotating search results on microblogging platforms with Wikipedia entities. Our evaluation has shown that a direct mapping of hashtags to a single dbpedia concept is not possible in many cases. One reason is the use of hashtags in different contexts, another reason is that hash tags do not necessarily describe an entity, but are rather used to describe contexts.

Another difficult problem we encountered is that different languages have to be dealt with: Topics are often commented on in a variety of languages on microblogging platforms. However, NLP techniques which are used to analyze the messages are language dependent.

%@@@ how to select external data (parse text from HTML, which part of web page to select - boilerplate stuff might spoil the results (reference blog cleansing paper

\bibliographystyle{abbrv}
\bibliography{bib}

\end{document}